---
title: "Framesoc Bench"
author: "Generoso Pagano"
date: "01/08/2015"
output:
  html_document: default
  pdf_document: default
---

```{r, echo=FALSE}
## RUN THIS BEFORE ALL

library("plyr")                                                                                                                                              
library("ggplot2")   

## Summarizes data.
## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
##   data: a data frame.
##   measurevar: the name of a column that contains the variable to be summariezed
##   groupvars: a vector containing names of columns that contain grouping variables
##   na.rm: a boolean that indicates whether to ignore NA's
##   conf.interval: the percent range of the confidence interval (default is 95%)
summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                      conf.interval=.95, .drop=TRUE) {
    require(plyr)

    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=FALSE) {
        if (na.rm) sum(!is.na(x))
        else       length(x)
    }

    # This does the summary. For each group's data frame, return a vector with
    # N, mean, and sd
    datac <- ddply(data, groupvars, .drop=.drop,
      .fun = function(xx, col) {
        c(N    = length2(xx[[col]], na.rm=na.rm),
          mean = mean   (xx[[col]], na.rm=na.rm),
          sd   = sd     (xx[[col]], na.rm=na.rm)
        )
      },
      measurevar
    )

    # Rename the "mean" column    
    datac <- rename(datac, c("mean" = measurevar))

    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean

    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult

    return(datac)
}

# utility to get trace data
get_trace <- function(data, s) {
  df <- data.frame(data)
  df <- df[df$size==s,]
  return(df)
}

```

# Intro

## Summary
Framesoc infrastructure benchmark, concerning trace import times and event reading times.

## Vocabulary
* Parameter: something inuencing the behavior of the system.
* Factor: parameter you decide to variate during experiment.
* Level: value given to a factor.
* Metric: what we want to measure

# Environment

Hard disk details, obtained with hdparm.

```
[generoso@generosohp ~]$ sudo hdparm -i /dev/sdb1

/dev/sdb1:

 Model=Hitachi HDS721010CLA630, FwRev=JP4OA41A, SerialNo=JP2940N03MSP5V
 Config={ HardSect NotMFM HdSw>15uSec Fixed DTR>10Mbs }
 RawCHS=16383/16/63, TrkSize=0, SectSize=0, ECCbytes=56
 BuffType=DualPortCache, BuffSize=29999kB, MaxMultSect=16, MultSect=16
 CurCHS=16383/16/63, CurSects=16514064, LBA=yes, LBAsects=1953525168
 IORDY=on/off, tPIO={min:120,w/IORDY:120}, tDMA={min:120,rec:120}
 PIO modes:  pio0 pio1 pio2 pio3 pio4 
 DMA modes:  mdma0 mdma1 mdma2 
 UDMA modes: udma0 udma1 udma2 udma3 udma4 udma5 *udma6 
 AdvancedPM=no WriteCache=enabled
 Drive conforms to: unknown:  ATA/ATAPI-2,3,4,5,6,7

```

System details, obtained with custom script.

```
[generoso@generosohp ~]$ ./read_configuration.sh 
# OS details
Linux generosohp.imag.fr 3.9.10-100.fc17.x86_64 #1 SMP Sun Jul 14 01:31:27 UTC 2013 x86_64 x86_64 x86_64 GNU/Linux
# HW details
Number of CPUs: 12
CPU information (all cpus are equal):
- model name : Intel(R) Xeon(R) CPU E5-1660 0 @ 3.30GHz
- cache size : 15360 KB
- hyperthreading : active
Scaling governor: performance
RAM: 16360588 kB
```

# Benchmark
    
## Trace Import

### Description

* Factors: 
    + Trace size: measured in # of events, directly translatable in the DB size (MB)
    + Indexing: boolean (are we indexing on timestamp or not?)
* Levels
    + Trace size:
        - small: 1 Mevent (66.3 MB withoud index, 81.2 MB with index)
        - medium: 10 Mevent (692.6 MB without index, 843.9 MB with index)
        - big: 100 Mevent (7.5 GB without index, 9.3 GB with index)
    + Indexing:
        - true
        - false
* Metric:
    + Import time
* Other parameters: 
    + fake traces created with the tool Temictli (available on https://github.com/soctrace-inria/framesoc.various)
    + the Temictli configuration file used is provided in this archive (./conf/temictli_bench.conf).
    + all the traces have 2 event parameters for each event (this has been chosen as an average value over many trace formats)
* Experiments:
    + small traces: 30 repetitions
    + medium traces: 10 repetitions
    + big traces: 5 repetitions
    + NOTE: as we will see below this number of repetition is largely enough to ensure very small confidence indervals.

### Analysis

Import time given the number of events and the indexing property.

```{r, echo=FALSE}                                                                                                                                              
# read data
df <- read.csv("../results/temictli_bench.log")
# aggregate
re <- ddply(df, c("size", "index"), summarise, N = length(time), mean = mean(time/1000), sd = sd(time/1000), se = sd / sqrt(N))
# todo test this
#re <- summarySE(df, measurevar="time", groupvars=c("size", "index"))
#re <- na.omit(re)
# plot                                                                                                                                                             
ggplot(re, aes(x=size, y=time, colour=index)) +                 
geom_line() +
geom_errorbar(aes(ymin=time-se, ymax=time+se), width=.1) +                                                                                               
geom_point() +                                                                                                                                                 
scale_y_continuous(name="time (s)") +                                                                                                                          
scale_x_continuous(name="# of events")        
```

Text representation of the above plot.

```{r, echo=FALSE}
re
```

## Trace reading

### Description

Note: for factors and levels already seen above (e.g. trace size), we omit the description,
since it is the same.

* Factors: 
    + Trace size
    + Indexing
    + Reading parameters: boolean
    + Reading intervall: the number of events per interval (all events, 100k events, ...)
* Levels
    + Trace size:
        - small
        - medium
        - big
    + Indexing and Reading parameters:
        - true
        - false
    + Interval
        - all events
        - 100k events (choosen as best compromise between speed and ui interactivity)
* Metric:
    + Total read time
    + Interval read time
    + Max used memory
* Other parameters: 
    + fake traces created above
    + the program used for reading (fr.inria.soctrace.framesoc.bench) is available on github:
    (https://github.com/soctrace-inria/framesoc.various)
* Experiments:
    + small traces: 30 repetitions
    + medium traces: 10 repetitions
    + big traces: 5 repetitions
    + TODO NOTE: as we will see below this number of repetition is largely enough to ensure very small confidence indervals.

### Analysis

```{r, echo=FALSE}

```


```{r, echo=FALSE}
# prepare data summaries
df <- read.csv("../results/reader_first.log")

# total time summary
dftime <- summarySE(df, measurevar="total_time", groupvars=c("size", "index", "param", "interval"))

# interval summary (for interval tests)
dfinterval <- summarySE(df, measurevar="interval_time", groupvars=c("size", "index", "param", "interval"))
dfinterval <- dfinterval[dfinterval$interval!='0',]
head(dftime)
head(dfinterval)

# read all, par, index
readall <- dftime[dftime$interval=='0',]
# read nopar, interval, index
readint <- dftime[dftime$param=='false',]

small_readall <- get_trace(readall, 1000000)
medium_readall <- get_trace(readall, 10000000)
big_readall <- get_trace(readall, 100000000)
small_readint <- get_trace(readint, 1000000)
medium_readint <- get_trace(readint, 10000000)
big_readint <- get_trace(readint, 100000000)
```

#### Total read time, with and without param, with and without index.

Small trace

```{r, echo=FALSE}
ggplot(data=small_readall, aes(x=index, y=total_time, fill=param)) + geom_bar(stat="identity", position=position_dodge()) +
geom_errorbar(aes(ymin=total_time-ci, ymax=total_time+ci), width=.1, position=position_dodge(.9)) +
scale_y_continuous(name="Total time (ms)")  
```

Medium trace

```{r, echo=FALSE}
ggplot(data=medium_readall, aes(x=index, y=total_time, fill=param)) + geom_bar(stat="identity", position=position_dodge()) +
geom_errorbar(aes(ymin=total_time-ci, ymax=total_time+ci), width=.1, position=position_dodge(.9)) +
scale_y_continuous(name="Total time (ms)")  
```

Big trace

```{r, echo=FALSE}
ggplot(data=big_readall, aes(x=index, y=total_time, fill=param)) + geom_bar(stat="identity", position=position_dodge()) +
geom_errorbar(aes(ymin=total_time-ci, ymax=total_time+ci), width=.1, position=position_dodge(.9)) +
scale_y_continuous(name="Total time (ms)")  
```

#### Total read time and interval time, without param, with and without index.
```{r, echo=FALSE}
ggplot(data=small_readint, aes(x=interval, y=total_time, fill=index) + 
geom_bar(stat="identity", position=position_dodge()) +
geom_errorbar(aes(ymin=total_time-ci, ymax=total_time+ci), width=.1, position=position_dodge(.9)) +
  scale_y_continuous(name="Total time (ms)")  
```

#### Total read time, with and without param, with and without index.

### Global analysis (obsolete)

Parse the summary with python script, to create configuration merging 3 factors (the output of parsing is in parsed.log).

```{r, echo=FALSE}
df <- data.frame(read.csv("../results/parsed.log"))
small <- df[df$size=='1000000',]
# plot                                                                                                                                                             
ggplot(data=small, aes(x=size, y=tmean, fill=config)) + geom_bar(stat="identity", position=position_dodge())                                                                                                                     
```
