---
title: "Framesoc Bench"
author: "Generoso Pagano"
date: "01/08/2015"
output: html_document
---

# Intro

## Summary
Framesoc infrastructure benchmark, concerning trace import times and event reading times.

## Vocabulary
* Parameter: something inuencing the behavior of the system.
* Factor: parameter you decide to variate during experiment.
* Level: value given to a factor.
* Metric: what we want to measure

# Environment

Hard disk details, obtained with hdparm.

```
[generoso@generosohp ~]$ sudo hdparm -i /dev/sdb1

/dev/sdb1:

 Model=Hitachi HDS721010CLA630, FwRev=JP4OA41A, SerialNo=JP2940N03MSP5V
 Config={ HardSect NotMFM HdSw>15uSec Fixed DTR>10Mbs }
 RawCHS=16383/16/63, TrkSize=0, SectSize=0, ECCbytes=56
 BuffType=DualPortCache, BuffSize=29999kB, MaxMultSect=16, MultSect=16
 CurCHS=16383/16/63, CurSects=16514064, LBA=yes, LBAsects=1953525168
 IORDY=on/off, tPIO={min:120,w/IORDY:120}, tDMA={min:120,rec:120}
 PIO modes:  pio0 pio1 pio2 pio3 pio4 
 DMA modes:  mdma0 mdma1 mdma2 
 UDMA modes: udma0 udma1 udma2 udma3 udma4 udma5 *udma6 
 AdvancedPM=no WriteCache=enabled
 Drive conforms to: unknown:  ATA/ATAPI-2,3,4,5,6,7

```

System details, obtained with custom script.

```
[generoso@generosohp ~]$ ./read_configuration.sh 
# OS details
Linux generosohp.imag.fr 3.9.10-100.fc17.x86_64 #1 SMP Sun Jul 14 01:31:27 UTC 2013 x86_64 x86_64 x86_64 GNU/Linux
# HW details
Number of CPUs: 12
CPU information (all cpus are equal):
- model name : Intel(R) Xeon(R) CPU E5-1660 0 @ 3.30GHz
- cache size : 15360 KB
- hyperthreading : active
Scaling governor: performance
RAM: 16360588 kB
```

# Benchmark
    
## Trace Import

* Factors: 
    + Trace size: measured in # of events, directly translatable in the DB size (MB)
    + Indexing: boolean (are we indexing on timestamp or not?)
* Levels
    + Trace size:
        - small: 1 Mevent (64 MB withoud index, 78 MB with index)
        - medium: 10 Mevent (661 MB without index, 805 MB with index)
        - big: 100 Mevent (7 GB without index, 8.7 GB with index)
    + Indexing:
        - true
        - false
* Metric:
    + Import time
* Other parameters: 
    + fake traces created with the tool Temictli (available on https://github.com/soctrace-inria/framesoc.various)
    + all the traces have 2 event parameters for each event (this has been chosen as an average value over many trace formats)

## Analysis

Import time given the number of events and the indexing property.

```{r, echo=FALSE}
library("data.table")                                                                                                                                              
library("ggplot2")                                                                                                                                                 
# create a summary: size, index, mean, error                                                                                                                        
dt <- data.table(read.csv("temictli_bench.log"))                                                                                      
setkey(dt,size,index)                                                                                                                                             
results = dt[,.(mean=mean(time/1000), sd=sd(time/1000), n=NROW(time/1000)), by="size,index"]                                                                                       
summary = data.frame(results[,.(size, index, mean, error = qnorm(0.975)*sd/sqrt(n))])                                                                               
# plot                                                                                                                                                             
ggplot(summary, aes(x=size, y=mean, colour=index)) +                                                                                                                
geom_errorbar(aes(ymin=mean-error, ymax=mean+error), width=.1) +                                                                                               
geom_point() +                                                                                                                                                 
scale_y_continuous(name="time (s)") +                                                                                                                          
scale_x_continuous(name="# of events")                                                                                                                          
```

Text representation of the above plot.

```{r, echo=FALSE}
summary
```

TODO: maybe a linear regression for each group (index, no index)